---
title: "Accessing MetaLab data"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Accessing MetaLab data}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
# all data chunks require network access, so they are skipped on CRAN
knitr::opts_chunk$set(
  collapse = TRUE, comment = "#>",
  fig.width = 6, fig.height = 4,
  eval = identical(Sys.getenv("NOT_CRAN"), "true")
)
```

[MetaLab](https://metalab.stanford.edu) is a database of community-augmented
meta-analyses of language acquisition and cognitive development: thousands of
standardized effect sizes, coded from the primary literature by dataset
curators, with the moderators needed to analyze them. metalabr reads MetaLab
data into R.

```{r attach}
library(metalabr)
```

## Released data

MetaLab data are published as versioned, citable releases (hosted on
[Redivis](https://stanford.redivis.com/datasets/81tq-8dp5ge6b9)).
`get_metalab_data()` reads a release — one row per effect size across all
datasets — and announces which release it used:

```{r get-data}
metalab_data <- get_metalab_data()
if (!is.null(metalab_data)) {
  dim(metalab_data)
}
```

For reproducible analyses, pin the release your paper used rather than
tracking `"current"`:

```{r pin-version}
metalab_2023 <- get_metalab_data(version = "2023.1")
if (!is.null(metalab_2023)) {
  nrow(metalab_2023)
}
```

`get_metalab_versions()` lists the available releases:

```{r versions}
str(get_metalab_versions(), max.level = 2)
```

Reading released data requires the `redivis` package (not on CRAN):

```{r install-redivis, eval = FALSE}
install.packages("redivis",
                 repos = c("https://langcog.r-universe.dev",
                           "https://cloud.r-project.org"))
```

## The dataset registry

`get_metalab_metadata()` reads the registry: one row per dataset, with
domains, citations, curators, summary counts, and each dataset's coded
moderators:

```{r metadata}
metadata <- get_metalab_metadata()
if (!is.null(metadata)) {
  metadata[1:5, c("name", "domain", "num_papers", "num_experiments")]
}
```

## Working with effect sizes

Individual datasets can be selected on read. Every dataset carries the
calculated effect sizes (`d_calc`, `g_calc`, `r_calc`, `log_odds_calc`),
their variances, and the standard MetaLab derived columns (`mean_age_months`,
`same_infant_calc` for clustering, and so on):

```{r mutex}
mutex <- get_metalab_data(short_names = "mutex", version = "2023.1")
if (!is.null(mutex)) {
  summary(mutex$g_calc)
}
```

The package includes the standard MetaLab visualizations, backed by the same
multilevel random-effects models (`metafor::rma.mv` with effect sizes nested
in participant groups nested in papers) used on the MetaLab site — scatter,
violin, forest, and funnel plots, plus Egger's regression test for funnel
asymmetry:

```{r funnel, fig.alt = "Funnel plot of the mutual exclusivity dataset"}
if (!is.null(mutex)) {
  metalab_funnel_plot(mutex, "mutex")
}
```

```{r funnel-test}
if (!is.null(mutex)) {
  metalab_funnel_test(mutex, "mutex")
}
```

## Current data without dependencies

`get_current_metalab_data()` downloads the current release's data bundle
from the MetaLab site with no account or extra packages, loading
`metalab_data` and `dataset_info` into the global environment (the legacy
metalabr interface, kept for backward compatibility):

```{r current}
get_current_metalab_data()
if (exists("dataset_info")) {
  nrow(dataset_info)
}
```

## Citing MetaLab

If you use MetaLab data, please cite the release you used (shown in every
data-access message), the dataset(s) you analyzed (citations are in the
registry's `full_citation` column), and the MetaLab platform papers — see
[metalab.stanford.edu](https://metalab.stanford.edu) for the current
citation policy.
